%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Introduction}  \newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%================================================================
\section*{A Message from Our Board}
%================================================================

\vspace{1em}
Dear Friends, Supporters, and Community Partners,

We are proud to present the inaugural annual fiscal report for the Imagination Library of San Antonio (ILSA). As a newly established 501(c)(3) nonprofit organization, 2024 marked a pivotal year in our journey to transform early childhood literacy in our beloved city.

\subsection*{Our Foundation}

In 2024, ILSA officially became a Local Program Partner of Dolly Parton's Imagination Library, joining a global network of communities committed to putting books in the hands of children from birth to age five. This milestone represents more than an organizational achievement---it signifies our community's collective commitment to ensuring that every child in San Antonio has access to the fundamental building blocks of learning and imagination.

\subsection*{Our Mission in Action}
At ILSA, we believe that books are the building blocks of a brighter future. Our mission is clear and unwavering: to ensure that all young children in San Antonio may have books to read. We achieve this by enrolling kids in Dolly Parton's Imagination Library. By providing free, high-quality books to children from birth to age 5, we support early childhood education and inspire a lifelong love of reading.

\subsection*{Our Commitment}
Guided by our small but dedicated board of community-minded individuals, we are driven by three fundamental goals:

\begin{itemize}[leftmargin=1em] 
  \item \textbf{Universal Enrollment}: Achieving equitable access to books by enrolling all eligible children in every zip code throughout San Antonio, Texas
  \item \textbf{Community Investment}: Increasing donations and sponsorships to launch, grow, and sustain this vital program for all of San Antonio's children
  \item \textbf{Educational Impact}: Enhancing local kindergarten readiness by investing in critical foundational reading skills
\end{itemize}

\subsection*{Looking Forward}
This report details our financial stewardship, operational achievements, and the generous support from donors and sponsors who share our vision. Every dollar invested in ILSA represents an investment in San Antonio's future---one book, one child, one brighter tomorrow at a time.
We invite you to explore the pages that follow to see how your support has translated into tangible impact for our community's youngest learners. Together, we are building not just literacy, but hope, opportunity, and brighter futures for all.

\vspace{1em}
With gratitude,
\vspace{2em}
\textbf{The Board of Directors}\
Imagination Library of San Antonio
\vspace{2em}
\rule{\textwidth}{1pt
}

\vspace{1em}
\noindent \textit{For questions about this report or our programs, please contact us at: \\
 \href{mailto:admin@imaginationlibrarysa.org}{admin@imaginationlibrarysa.org}}

\vspace{0.5em}

%\noindent \textbf{Connect with us:}\\
%Facebook ||
%| Instagram ||
%| LinkedIn\\

\noindent Learn more about Dolly Parton's Imagination Library: \href{https://imaginationlibrary.com/usa/}{imaginationlibrary.com/usa}